---
title: "Kaggle2018"
author: "Kuldeep Singh Chouhan"
date: "26/10/2019"
output: 
  html_document:
    number_section: false
    fig_caption: true
    toc: yes
    theme: cosmo
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE)
```


# **Introduction**

This is Kaggle's *2018 Kaggle Machine Learning and Data Science Survey*. KAggle conducted an industry-wide survey that presents a truly comprehensive view of the state of data science and machine learning. In this kernel I will perform extensive EDA to get interesting stories from the data.

# **Loading data and libraries** {.tabset .tabset-fade .tabset-pills}

We will start our EDA by checking the dimensions of our dataset. The data contains following:

## Loading libraries

```{r message=FALSE}
#loading libraries

library(tidyverse)
library(ggthemes)
library(data.table)
```

## Data

```{r}
multi <- fread("data/multipleChoiceResponses.csv", skip = 1)

free <- fread("data/freeFormResponses.csv", skip = 1)
```

# **Demmographics**

Let's examine demographic data. We will first look at gender.

## Gender

We will see the gender distribution of respondents.

```{r}

vars <- c(gender = "What is your gender? - Selected Choice",
          gen_txt = "What is your gender? - Prefer to self-describe - Text",
          country = "In which country do you currently reside?",
          age = "What is your age (# years)?")

multi <- multi %>%
  rename(!!vars)

```

```{r}
gen <- multi %>%
  group_by(gender)%>%
  count()

gen %>%
  ggplot(aes(gender, n, fill = gender)) +
  geom_col() +
  geom_label(aes(label = n)) +
  labs(x = "Gender", y = "Respondents") +
  coord_flip() +
  theme_gdocs() +
  theme(legend.position = "none") +
  ggtitle("Gender Counts")

```

## Age

Now we will see the age distribution of respondents.

```{r}
multi <- multi %>%
  mutate(age = fct_relevel(age, c("18-21", "22-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-69","70-79","80+")))  

ag <- multi %>%
  group_by(age) %>%
  count() 

ag %>%
  ggplot(aes(fct_reorder(age, n), n, fill = age)) +
  geom_col() +
  geom_label(aes(label = n)) +
  labs(x = "Age Group", y = "Respondents Counts") +
  coord_flip() +
  theme_gdocs() +
  theme(legend.position = "none") +
  ggtitle("Age Groups")

```


